---
title: "Does Transmission Type affect Mileage?"
author: "Wang Xin"
date: "2016-4-25"
output: pdf_document
---

## Executive Summary

In responding of the insterests of _Motor Trend_, a magazine about the automobile industry, this report explores the relationship between a set of variables and mile per gallon(MPG), trying to answer two questions:

* Is an automatic or manual transmission better for MPG
* Quantify the MPG difference between automatic and manual transmissions

__Notice__: This work is part of Coursera regression model courese's assignment.  The results could be wrong or outdated.  

## Load and Prepare Data

```{r}
data("mtcars")
dt <- mtcars
dt$am<-factor(dt$am,levels=c(0,1),labels=c('Automatic','Manual'))
```

## Exploratory Data Analyse #1

By simply boxplot mpg by different transmission type (See Appendix #1), it shows that manual got higher mpg than automatic. 

## Building Model #1

```{r, eval=FALSE}
fit1 <- lm(mpg ~ am, data = dt)
summary(fit1)
```
It shows that an automatic car with 17.147 mpg, gets 7.245 mpg more if it is manual transimission.  The Adjusted R-squared vaue is 0.3385, which meacs the model explain only 34% of the MPG varibles. So, more variables need to be indroduced into the model. 

## ANOVA test

Implement ANOVA test to determine those most relevant variables.

```{r}
dt_anova <- aov(formula = mpg ~ ., data = dt)
summary(dt_anova)
```

The results suggest that disp(displacement), wt(Weight) and cyl(Number of cylinders) are those top 3 significant variables.

## Exploratory Data Analyse #2

Visiualize more variables (See appendix #2) and it shows that disp,wt,cyl all have stronger relationship with mpg. 

## Building Model #2

Build new model with the other 3 varibales:

```{r}
fit2 <- lm(mpg ~ cyl + disp + wt + am, data = dt)
summary(fit2)
```

It shows that disp is not significant for the model.  So I drop this varible.

## Building Model #3

```{r}
fit3 <- lm(mpg ~ cyl + wt + am, data = dt)
summary(fit3)
```
This model has Residual standard error as 2.612 on 28 degree of freedom, and Adjusted R-squared value os 0.8122, which means it can explain 81% of the MPG variables. All of the coeddicients are significant at 0.05 significant level.  This is a good one.

## Residual plot and diagnostics

According to the residual plots (See Appendix #3):

1. The residuals vs fitted plot shows no consistent pattern
2. The points lie closely to the line, so the residuals are normally distributed
3. The scale-location plot confirms the constant variance assumption
4. No outliers are present. 

## Conclusion

In this dataset, although manual vehicles achieve more mpg that automatics, thransmission type is more a good factor to predict mpg. The number of cylinders and the gross weight are much more directly linked to fuel usage.




## Appendix #1 - botplot
```{r, fig.width=5, fig.height=3}
plot(mpg ~ am, data = dt, main = "Miles/Gallon vs. Transmission Tpye")
```

## Appendix #2 - plot more variables
```{r,fig.height=3}
library(ggplot2)
qplot(x = mpg, y = disp, data = dt,
      size = wt, color = factor(cyl), facets=.~am,
      main='Variables Effect on MPG')
```

## Appendix #3 - residual plots
```{r, fig.height=8}
par(mfrow = c(2,2))
plot(fit3)
```